\documentclass[10pt,final,conference,comsoc]{IEEEtran}

% reviewer 1, comment 1 and 7 haven't been extended.
% Some figure maybe need to be fixed a little bit.
\usepackage[utf8]{inputenc}

\usepackage[T1]{fontenc} % optional
\usepackage{amsmath,amssymb,amsfonts}
\usepackage[cmintegrals]{newtxmath}
\usepackage{cite}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{bm}
\usepackage{siunitx}
\usepackage{comment}
\usepackage{mathtools}
\usepackage{url}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\newcommand{\LRT}[2]{
  \mathrel{\mathop\gtrless\limits^{#1}_{#2}}
}
\newcommand{\numb}[1]{\SI{#1}}
\newcommand{\dB}{\decibel}
\newcommand{\est}[2]{\hat{#1}_{#2}}
\newcommand{\sd}{\text{SD}}
\newcommand{\cn}{\mathcal{CN}}
\newcommand{\n}{\mathcal{N}}
\newcommand{\opt}{\text{opt}}
\newcommand{\nm}{\text{NM}}
\newcommand{\J}{\mathrm{J}}
\newcommand{\prob}[1]{\text{P}_{#1}}
\newcommand{\fa}{\text{FA}}
\newcommand{\D}{\text{D}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Q}{\mathrm{Q}}
\newcommand{\Lagr}{\mathcal{L}}


\graphicspath{{../main_matlab_figures/}{../matlab/}}

\DeclareUnicodeCharacter{0303}{}

\begin{document}

\title{Low Complexity Methods for Joint Detection and Synchronization of TDMA Bursts
}

\author{\IEEEauthorblockN{Haotian Zhai 
  and Bernd-Peter Paris}
\IEEEauthorblockA{\textit{Department of Electrical and Computer Engineering} \\
\textit{George Mason University}\\
Fairfax, VA 22030 \\
\{hzhai,pparis\}@gmu.edu}
}

\maketitle


\begin{abstract}

This paper proposes a data-aided joint detection and synchronization algorithm for TDMA bursts.
A sequential detection algorithm based on the Generalized Likelihood Ratio Test (GLRT) is used to 
detect the embedded preamble signal in received data stream. 
% sequential detection algorithm is necessary to be derived.
Carrier synchronization is attempted for each sample instant during sequential detection and
resulting phase and frequency estimates are used by the GLRT. 
To make this algorithm implemented in software-defined radio (SDR),
a low complexity carrier estimation algorithm with good estimation accuracy is proposed.
Then, a refinement of the carrier estimate is computed when the preamble has been detected.
% SDR is also mentioned in abstract.
It is shown that this estimate approaches the Cramer-Rao bound even at low SNR. 
The complete joint detection and estimation procedure is validated through simulations and SDR.

\end{abstract}

\begin{IEEEkeywords}
Joint detection and estimation, GLRT, low computational complexity, Cramer-Rao vector bound, Software-defined radio.
\end{IEEEkeywords}

\input{introduction}

\input{model}

\input{detection}

\input{freq_est}

\input{simulations}

\input{real_implementation}

% % extension to reviewer 1, comment 8
% \input{conclusion} can be finished later

% extension to reviewer 1, comment 6, reviewer 3, comment 4
\input{appendices}

\bibliographystyle{ieeetr}
\bibliography{ref.bib}

\end{document}


