\section{Frequency and Phase Estimation}%
\label{sec:freq_est}   

In this section, estimation is performed using a window of $N$ samples, $0\leq n<N$, starting
at the position $\bar{p}$ of the preamble.
For estimating the frequency offset $\delta$ and phasor $\xi=Ae^{j\phi}$, the maximum likelihood estimation (MLE) of the parameters in~\eqref{eq:model} is given by

\begin{equation}
    \label{eq:ML_f_xi}
      \hat{\delta},\hat{\xi}=\min_{\delta,\xi=Ae^{j\phi}}\sum_{n=0}^{N-1}|r_n-s_n\xi e^{j2\pi\delta n}|^{2}.
    \end{equation}
A closed form for $\hat{\xi}$ is readily derived by taking the Wirtinger derivative with $\xi$ and setting it equal to zero,

\begin{equation}
    \label{eq:opt_xi}
    \hat{\xi}=\frac{\sum_{n=0}^{N-1}{r_{n}s_n^{*}e^{-j2\pi\hat{\delta} n}}}{\sum_{n=0}^{N-1}|s_{n}|^2},
  \end{equation}
so that $\hat{\xi}$ relies on the frequency estimate $\hat{\delta}$, and $\hat{\phi}=\arg\{\hat{\xi}\}$. 
% Moreover, by plugging~\eqref{eq:opt_xi} into~\eqref{eq:generalized_corr}, the GLRT based detector can be computed more
% efficiently,

% \begin{equation}
%     \label{eq:simplified_GLRT_detector}
%     \rho(p{+}\Delta) \approx
%     \frac{|\tilde{\xi}_{p+\Delta}|}
%     {||\bm{r}_{p+\Delta}||\cdot||\bm{s}||} \LRT{H_1}{H_0} \gamma
%   \end{equation}
% where $\tilde{\xi}_{p+\Delta}=\sum_{n=0}^{N-1}{r_{n+p+\Delta}s_n^{*}e^{-j2\pi\hat{\delta}_{p+\Delta} n}}$ represents 
% the cross-correlation (the numerator) of the phasor estimate $\hat{\xi}$ in~\eqref{eq:opt_xi} at position $p{+}\Delta$, and $||\bm{s}||$ is the Euclidean norm of the preamble.

A necessary condition for the frequency offset estimate $\hat{\delta}$
is obtained similarly by taking the derivative of~\eqref{eq:ML_f_xi} with respect to
$\delta$ and setting it equal to zero. Skipping all intermediate derivation steps for brevity's sake,
it yields

\begin{equation}
    \label{eq:necessary condition for delta}
    J(\hat{\delta}) = \Im\bigg\{\sum_{k=1}^{N-1}{\sum_{m=k}^{N-1}{kr_{m-k}r_m^{*}s_{m-k}^{*}s_m}e^{j2\pi\hat{\delta}k}}\bigg\}=0.
    \end{equation}

There are a number of local minima of~\eqref{eq:ML_f_xi} also satisfying
the necessary condition for $\hat{\delta}$ of~\eqref{eq:necessary condition for delta} in addition to the absolute minimum
(the exact solution of MLE). In~\cite{Luise_Reggiannini_95} and~\cite{Fitz_94}, the "false minima" are avoided
by appropriately restricting the operating range of the estimator. Specifically,
instead of calculating the sample autocorrelation functions for all lags $k\in[1,N-1]$, 
they truncate~\eqref{eq:necessary condition for delta} by only considering the lag autocorrelation functions for $k{\in}[1,\varepsilon]$,
where $\varepsilon \ll N{-}1$. This is to avoid 
including the inner summations in~\eqref{eq:necessary condition for delta} with few
terms and correspondingly high variance.

Moreover, the estimator $\hat{\delta}$ in~\eqref{eq:necessary condition for delta} has no closed-form solution.
In~\cite{Luise_Reggiannini_95}, the necessary condition is approximated by replacing the exponential with its Taylor series expansion.
In~\cite{Fitz_94}, an approximate solution is obtained via Euler's identity for large $N$.
Both L\&R~\cite{Luise_Reggiannini_95} and Fitz~\cite{Fitz_94} estimators have computational complexity $O(N^2)$ 
reflecting the double summation. In~\cite{kay_89}, the Kay estimator reduces the complexity from $O(N^2)$ to $O(N)$ by only computing~\eqref{eq:necessary condition for delta} at lag $k=1$.
However, it suffers from poor accuracy at low SNRs.

In this paper, we propose a family of alternative solutions to~\eqref{eq:necessary condition for delta}.
A coarse solution with $O(N)$ complexity is used for operating at high sample rate during the sequential detection.
It prioritizes low complexity at the expense of some loss of accuracy. A second more accurate solution is used to improve
the estimation accuracy at moderate complexity to enable co-herent demodulation once the preamble has been detected. 

\subsection{Coarse Solution: Single-Lag Estimator with Length-v Partial Correlating}

The first estimator is rooted in the insight that at high SNR, every lag $k$ in~\eqref{eq:necessary condition for delta}
can be used to approximate the frequency estimate $\delta$. By setting $r_m\approx s_m\xi e^{j2\pi\delta m}$,
\eqref{eq:necessary condition for delta} is expanded to

\begin{equation}
  \label{eq:delta_extens_no_noise}
  \Im\bigg\{A^2\sum_{k=1}^{N-1}\sum_{m=k}^{N-1}k|s_{m-k}|^2|s_m|^2e^{j2\pi (\hat{\delta}-\delta)k}\bigg\}=0.
  \end{equation}
Note, the inner summation in~\eqref{eq:delta_extens_no_noise} is purely real for every lag $k$ if $\hat{\delta}=\delta$.
This suggests that an unbiased estimate of $\delta$ can be obtained by using only a single lag $k$. The approach yields
a closed-form solution for $\hat{\delta}$, which is given by

\begin{equation}
  \label{eq:single_lag_estimator_wout_partial_corr}
  \hat{\delta}_{SL}(k)=-\frac{\arg\big\{\sum_{m=k}^{N-1}r_{m-k}r_m^*s_{m-k}^*s_m\big\}}{2\pi k},
\end{equation}
However, the above single-lag (SL) estimator has insufficient
accuracy at low SNRs as we will show below. To extend range of operation to low SNR,
an alternative SL estimator including partial coherent integrator
prior to estimation is given by

\begin{equation}
  \label{eq:single_lag_estimator_w_partial_corr}
  \hat{\delta}_{SL}^{(v)}(k_v)=-\frac{\arg\big\{\sum_{l=k_v}^{N/v-1}\digamma_l^*\digamma_{l-k_v}\big\}}{2\pi k_vv},
\end{equation}
where $\digamma$ denotes the result of coherent integration of length-$v$ blocks

\begin{equation}
  \label{eq:coherent_integrator}
  \digamma_l=\sum_{n=lv}^{(l+1)v-1}r_ns_n^*, \quad \text{for}~l=0,1,\ldots,N/v{-}1.
\end{equation}
Normally, $v$ is set to be a factor of $N$ to include all the sample instants with maximum $v=N/2$.
In~\eqref{eq:single_lag_estimator_w_partial_corr}, $k_v=\lfloor k/v \rfloor$ denotes
the distance between blocks that we want to estimate carriers from.
$\lfloor \cdot \rfloor$ is the floor operator. It can be seen 
\eqref{eq:single_lag_estimator_wout_partial_corr} is the special case of~\eqref{eq:single_lag_estimator_w_partial_corr}
when $v=1$, meaning no partial integrator is used.


% Next, we will show that by using a so-called length-$v$ coherent integrator,
% the single-lag estimator also gains a good accuracy at low SNRs.

% \subsubsection{Closed-form expression}
% The single-lag (SL) estimator with length-$v$ partial correlating is given by

% \begin{equation}
%   \label{eq:single_lag_estimator_w_partial_corr}
%   \hat{\delta}_{SL}^{(v)}(k_v)=-\frac{\arg\big\{\sum_{l=k_v}^{N/v-1}\digamma_l^*\digamma_{l-k_v}\big\}}{2\pi k_vv},
% \end{equation}
% where $\digamma$ denotes the coherent integrator, and

% \begin{equation}
%   \label{eq:coherent_integrator}
%   \digamma_l=\sum_{n=lv}^{(l+1)v-1}r_ns_n^*, \quad \text{for}~l=0,1,\ldots,N/v{-}1.
% \end{equation}
% In~\eqref{eq:coherent_integrator}, $v$ is the number of coherent correlations for averaging in each coherent integrator. 
% Normally, $v$ is set to be a factor of $N$ to include all the sample instants.
% In~\eqref{eq:single_lag_estimator_w_partial_corr}, $k_v$ denotes the distance between
% two coherent integrators for calculating the frequency estimate from non-coherent sample instants;
% $k_v=\lfloor k/v \rfloor$, where $\lfloor \cdot \rfloor$ is the floor operation. 
% Particularly, when $v{=}1$, meaning no partial correlating is used,~\eqref{eq:single_lag_estimator_w_partial_corr} reduces to

% \begin{equation}
%   \label{eq:single_lag_estimator_wout_partial_corr}
%   \hat{\delta}_{SL}^{(1)}(k)=-\frac{\arg\big\{\sum_{m=k}^{N-1}r_{m-k}r_m^*s_{m-k}^*s_m\big\}}{2\pi k},
% \end{equation}

\subsubsection{Performance of single-lag estimator}

For evaluating the performance of the SL estimator,
we first look at the probability density function (pdf)
for the two coherent integrations in~\eqref{eq:single_lag_estimator_w_partial_corr}.
By setting $r_n=s_n\xi e^{j2\pi \delta n}+w_n$ into~\eqref{eq:coherent_integrator}, each of
the two coherent integrators in~\eqref{eq:single_lag_estimator_w_partial_corr} yields a complex Gaussian
random variable (r.v.) with respective pdfs

\begin{equation}
  \label{eq:pdf_co_integrator}
  \begin{aligned}
    &\digamma_l^* \sim \cn\bigg(\xi^*\sum_{n=lv}^{(l+1)v-1}e^{-j2\pi \delta n},\frac{N_0}{2}v\bigg), \\
    &\digamma_{l-k_v} \sim \cn\bigg(\xi\sum_{n=(l-k_v)v}^{(l-k_v+1)v-1}e^{j2\pi \delta n},\frac{N_0}{2}v\bigg).
  \end{aligned}
\end{equation}
where $A^2|s_n|^2 \approx A^2= E_s/M$ is hold by assuming a constant and normalized envelope of the preamble given in Section~\ref{sec:model}.
Note, $\digamma_l^*$ and $\digamma_{l-k_v}$ are uncorrelated. Based on~\eqref{eq:pdf_co_integrator}, 
the product $C_{\digamma_{l}}=\digamma_l^*\digamma_{l-k_v}$ has a mixed distribution of a complex Gaussian and a
second kind Bessel function from the products of noise terms.
The mean $\mu_{C_{\digamma_{l}}}$ and variance $\sigma^2_{C_{\digamma_{l}}}$ are given by, 

\begin{equation}
  \begin{aligned}
  \label{eq:mean_var_product_coherent_int}
  \mu_{C_{\digamma_l}}&=\frac{E_s}{M}\sum_{n=lv}^{(l+1)v-1}e^{-j2\pi \delta n}\bigg(\sum_{m=(l-k_v)v}^{(l-k_v+1)v-1}e^{j2\pi \delta m}\bigg) \\
  &=\frac{E_s}{M}e^{-j2\pi \delta k_vv}\bigg(\frac{\sin(\pi \delta v)}{\sin(\pi \delta)}\bigg)^2, \\
  \sigma^2_{C_{\digamma_l}}&={\underbrace{\frac{N_0^2}{4}v^2}_{\text{from Bessel}}}+{\underbrace{2v\frac{N_0E_s}{2M}\bigg(\frac{\sin(\pi \delta v)}{\sin(\pi \delta)}\bigg)^2}_{\text{from Complex Gaussian}}},
  \end{aligned}
\end{equation}
where $\frac{\sin(\pi \delta v)}{\sin(\pi \delta)}$ is the Dirichlet function of $\delta$, which approaches
maximum value $v$ at $\delta{=}0$ and first two zeros at $\delta{=}\pm 1/v$.
Note, both $\mu_{C_{\digamma_l}}$, $\sigma^2_{C_{\digamma_l}}$ are independent of the partial integration block.
Thus, $\sum C_{\digamma_l}$ in the argument operator of~\eqref{eq:single_lag_estimator_w_partial_corr} 
also has the mixed distribution with a complex Gaussian and a second kind Bessel function.

Recall from~\eqref{eq:single_lag_estimator_w_partial_corr}, the distribution of SL estimator depends on $\arg\{\cdot\}$.
It can be shown that the full pdf of $\arg\{\zeta\}$, where $\zeta$ is complex Gaussian distributed, is well approximated for moderate SNR, as Gaussian.
Specifically,

\begin{equation}
  \label{eq:pdf_arg_comp_Gaus}
  \arg\{\zeta\} \sim \n\big(\angle \mu_{\zeta},\sigma^2_{\zeta}/|\mu_{\zeta}|^2\big).
\end{equation}
The derivation of~\eqref{eq:pdf_arg_comp_Gaus} is omitted due to space constraint.
Based on~\eqref{eq:pdf_arg_comp_Gaus}, the performance evaluation of 
the SL estimator can be obtained by looking at the variance in~\eqref{eq:pdf_arg_comp_Gaus},
or equivalently, the "output" SNR, which equals to the absolute square of mean to variance of
$\sum C_{\digamma_l}$ in the argument operator of~\eqref{eq:single_lag_estimator_w_partial_corr}, i.e.,

\begin{equation}
  \begin{aligned}
    \label{eq:SNR_out}
    \text{SNR}_{out}=\frac{|\mu_{\sum C_{\digamma_l}}|^2}{\sigma^2_{\sum C_{\digamma_l}}} 
    % &=\frac{\frac{Es^4}{M^4A^4}\Big(\frac{\sin(\pi \delta v)}{\sin(\pi \delta)}\Big)^4(N/v-k_v)^2}
    % {\Big[v^2\frac{N_0^2E_s^2}{4M^2A^4}+2v\frac{N_0E_s^3}{2M^3A^4}\Big(\frac{\sin(\pi \delta v)}{\sin(\pi \delta)}\Big)^2\Big](N/v{-}k_v)} \\
    =\frac{(N/v-k_v)\Big(\frac{\sin(\pi \delta v)}{\sin(\pi \delta)}\Big)^4}
    {\frac{v^2}{\text{SNR}_{in}}+2v\Big(\frac{\sin(\pi \delta v)}{\sin(\pi \delta)}\Big)^2}\cdot\text{SNR}_{in}. \\
  \end{aligned}
\end{equation}
where $\text{SNR}_{in}=\frac{2E_s}{MN_0}$. From~\eqref{eq:SNR_out},
we can see that the "output" SNR is degraded by the variance of second
Bessel r.v. at low input SNRs. By assuming $|\delta|v{\ll} 1$, the processing gain
of the SL estimator with length-$v$ partial correlator relative to the SL estimator
without partial correlator is obtained by the ratio of~\eqref{eq:SNR_out} 
with $v$ to~\eqref{eq:SNR_out} with $v=1$, which yields 

\begin{equation}
  \begin{aligned}
    \label{eq:relative_processing_gain}
    \text{Relative processing gain at small}~|\delta|v \approx\frac{v+2v\text{SNR}_{in}}{1+2v\text{SNR}_{in}}.
  \end{aligned}
\end{equation}

Some observations can be obtained from~\eqref{eq:relative_processing_gain}.
First, we see the maximum relative processing gain approaches to $v$ if there is no signal. 
Moreover, when the input SNR is fixed and small, the relative processing gain
is increased as $v$ becomes larger. On the other hand, at high input SNRs,
the effect of Derichlet function in~\eqref{eq:SNR_out} with respect to the frequency offset
$|\delta|v$ be-comes more important than the relative processing gain.
Thus, the $\text{SNR}_{out}$ approaches the maximum when $v=1$.

Based on~\eqref{eq:single_lag_estimator_wout_partial_corr},~\eqref{eq:mean_var_product_coherent_int}
and~\eqref{eq:pdf_arg_comp_Gaus}, the distribution of $\hat{\delta}_{SL}^{(v)}(k_v)$ at moderate SNRs is finally given by

\begin{equation}
  \label{eq:lower_bound_single_lag_high_snr}
  \hat{\delta}_{SL}^{(v)}(k_v) \sim \n \Bigg(\delta,\frac{(Mv+4\Big(\frac{\sin(\pi \delta v)}{\sin(\pi \delta)}\Big)^2E_s/N_0)\cdot Mv}{16\pi^2k^2(N/v{-}k_v)(E_s/N_0)^2\Big(\frac{\sin(\pi \delta v)}{\sin(\pi \delta)}\Big)^4}\Bigg).
\end{equation}
Thus, $\hat{\delta}_{SL}^{(v)}(k_v)$ is unbiased. 
A lower bound for the variance of the SL estimator at high SNRs is obtained
as the variance of~\eqref{eq:lower_bound_single_lag_high_snr} when $v=1$;
On the other hand, an approximate lower bound at low SNRs can be obtained by the variance of~\eqref{eq:lower_bound_single_lag_high_snr}
at maximum $v=N/2$.
Moreover, we see the variance also depends on
the value of $k_v$. The best choice for $k_v$ is to choose 
$k_v=\lfloor\frac{2N}{3v}\rfloor$ to minimize the variance.

\subsubsection{Estimation range}
The SL estimator may suffer an effect
of "aliasing" akin to $2\pi |\delta|k_vv{>}\pi$. Thus, a safe estimation range for the estimator with optimal $k_v=\lfloor\frac{2N}{3v}\rfloor$ to avoid modulo-$2\pi$ ambiguity
is $\delta$ within $\pm 3/(4MN)$. Compared with the same autocorrelation-based estimator, e.g.,
the L\&R~\cite{Luise_Reggiannini_95} and Fitz~\cite{Fitz_94}, single-lag estimator has 
$3/8$ estimation range of L\&R and $3/4$ estimation range of Fitz.

\subsubsection{Computational complexity}

We have discussed the accuracy of single-lag estimator.
It is also necessary to address the complexity since the SL estimator
is used in high sample-rate case. The computational complexity of single-lag estimator 
can be readily assessed from~\eqref{eq:single_lag_estimator_w_partial_corr},~\eqref{eq:coherent_integrator} and~\eqref{eq:single_lag_estimator_wout_partial_corr}.

Specifically, we compare with the complexity of SL estimators in sequential detection with and without partial correlating.
Note, without partial correlating, from~\eqref{eq:single_lag_estimator_wout_partial_corr},
$s_{m-k}^*s_m$ can be precomputed and stored
like "filter coefficients"; Moreover, due to the characteristic of the sequential detection process, the products of received samples, $r_{m-k}r_m^*$,
can be stored in a shift register so that only one new product needs to be computed per sample period.

The exact computational complexity of two single-lag estimators are given in Table~\ref{table:computational complexity comparison} with optimal $k_v \approx 2N/(3v)$. 
We see $\hat{\delta}_{SL}^{(1)}(k)$ has approximate 2 times fewer complex products and additions compared with $\hat{\delta}_{SL}^{(v>1)}(k_v)$ in sequential detection.
Furthermore, note the complexity of Kay estimator in~\cite{Morelli_Mengali_98} is given approximately $3N/4$ complex products and additions, 
which is slightly larger than $\hat{\delta}_{SL}^{(1)}(k)$.

\begin{table}[t]
  \caption{Complexity of single-lag estimators with and without partial correlating in sequential detection with optimal $k_v\approx2N/(3v)$}  % title of Table
  \centering 
  \begin{tabular}{c c c} 
  \hline\hline 
   & Complex products & Complex additions \\ [0.5ex] 
  \hline 
  $\hat{\delta}_{SL}^{(1)}(k)$  & $N/3+1$ & $N/3$ \\ 
  $\hat{\delta}_{SL}^{(v)}(k_v)$ & $(2+1/v) N/3$ & $(2-1/v)N/3-1$ \\ [1ex]
  \hline
  \end{tabular}
  \label{table:computational complexity comparison}
\end{table}

\subsection{Fine Solution: Newton-Method based Estimator}

The SL estimator emphasizes low-complexity property and is intended
to provide merely sufficient good carrier synchronization to enable coherent detection.
Once the signal has been acquired, the SL estimator can be improved by 
investing additional computations. Since detection events are rare, the computational
complexity is of little concern.

The principle of Newton-Method based estimator is to use the single-lag estimator as the starting point for a Newton-type iteration 
aimed at finding a better solution to the necessary condition~\eqref{eq:necessary condition for delta}. 
In principle, multiple iterations are possible to produce successively better approximations to the root of
$J'(\cdot)$ in~\eqref{eq:necessary condition for delta}. Specifically, the iterations are given by

\begin{equation}
  \label{eq:iter_NM_est}
  \hat{\delta}_{NM}^{(i+1)}=\hat{\delta}_{NM}^{(i)}-
  \frac{J(\hat{\delta}_{NM}^{(i)})}{J^\prime(\hat{\delta}_{NM}^{(i)})}
\end{equation}
where $\hat{\delta}_{NM}^{(0)}=\hat{\delta}^{(v)}_{SL}(k_v)$ is the starting point of iteration and
$J^\prime(\cdot)$ denotes the derivative of $J$ with respect to $\hat{\delta}$. 
Our simu-lations indicate that only a single iteration is usually sufficient to achieve very good accuracy.

From~\eqref{eq:iter_NM_est} and the previous discussion, we can conclude the importance of accuracy of the SL estimator
at low SNRs: 
with a merely sufficient good accuracy, the SL estimator not only increases the probability of detection by better
fitting the preamble and received signal as in sequential detector~\eqref{eq:generalized_corr}, but 
it provides a reasonable starting point for getting the more accurate NM estimator. In simluations,
we will also show the case when the NM estimator has a worse accuracy than the SL estimator if the latter does not provide enough accuracy. 

% ambiguity problem for sl.